---
title: "The What, Why, and How of Interactive Visualizations"
output: html_notebook
runtime: shiny
---

# The What, Why, and How of Interactive Visualizations

## What?

Check out [examples](https://shiny.rstudio.com/gallery/).

## Why?

Benefits

-   Fitting more information into a smaller space (expressive)

-   Letting users adapt data visualization to their needs

-   Facilitating collaboration between data analysts/scientists and less data-savvy people

See [The Surprising Story Behind The Mac's Original Calculator App](https://www.businessinsider.com/macintosh-calculator-2011-10).

![](https://i.insider.com/4ea5b2d869bedd576a000032?width=300&format=jpeg&auto=webp)

## How?

Use [shiny](https://shiny.rstudio.com/tutorial/) (see yaml header).

This example uses the [Palmer Penguins Dataset](https://allisonhorst.github.io/palmerpenguins/).

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(shiny)
library(tidyverse)
library(palmerpenguins)
```

```{r}
penguins_df <- penguins %>% drop_na()
penguins_df
```

### Categorical Input

```{r}
penguins_df %>% select(species) %>% unique()
```

```{r}
penguins_df %>%
  filter(species == "Chinstrap") %>% 
  ggplot() + 
    geom_point(
        mapping = aes(
          x = bill_length_mm, 
          y = bill_depth_mm,
          color = species
       )
    ) + 
    xlim(30, 60) + 
    ylim(15, 22)
```

In this example, we'll use a `checkboxGroupInput` to look at specific species of penguins.

    checkboxGroupInput(
      inputId = ..., # Think of it as an object name (not visible to users)
      label = ..., # Title (visible to users)
      choices = ..., # Options for users to select
      selected = ... # What's selected by default
    )

Use `renderPlot(…)` and `input$inputId` to generate your interactive viz.

    renderPlot(...)

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

checkboxGroupInput(
  inputId = "species_selection", # Think of it as an object name (not visible to users)
  label = "Penguin species:", # Title (visible to users)
  choices = c("Adelie", "Chinstrap", "Gentoo"), # Options for users to select
  selected = c("Adelie", "Chinstrap", "Gentoo") # What's selected by default
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

renderPlot(
  penguins_df %>%
    filter(species == input$species_selection) %>% 
    ggplot() + 
      geom_point(
          mapping = aes(
            x = bill_length_mm, 
            y = bill_depth_mm,
            color = species
         )
      )
)
```

Other tips

-   Set the `xlim` and `ylim`

### Numerical Variables

In this example, we'll use a `sliderInput` to do a sample of the penguins (`sample_frac`).

```{r}
penguins_df %>%
  sample_frac(size = 0.75) %>% 
  ggplot() + 
    geom_point(
        mapping = aes(
          x = bill_length_mm, 
          y = bill_depth_mm,
          color = species
       )
    ) + 
    xlim(30, 60) + 
    ylim(15, 22)
```

    sliderInput(
      inputId = ..., # Think of it as an object name (not visible to users)
      label = ..., # Title (visible to users)
      min = ..., # Minimum value on slider
      max = ..., # Maximum value on slider
      value = ..., # Default value on slider
      step = ... # Interval between each tick on the slider
    )

User `renderPlot(…)` and `input$inputId` to generate your interactive viz.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

sliderInput(
  inputId = "sample_size", # Think of it as an object name (not visible to users)
  label = "Sample Size:", # Title (visible to users)
  min = 0, # Minimum value on slider
  max = 1, # Maximum value on slider
  value = 1, # Default value on slider
  step = 0.01 # Interval between each tick on the slider
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

renderPlot(
  penguins_df %>%
    sample_frac(size = input$sample_size) %>% 
    ggplot() + 
      geom_point(
          mapping = aes(
            x = bill_length_mm, 
            y = bill_depth_mm,
            color = species
         )
      ) + 
      xlim(30, 60) + 
      ylim(15, 22)
)
```

### Multiple Inputs

Make sure you have unique `InputId`s!

```{r}
penguins_df %>%
  filter(species == "Adelie") %>% 
  sample_frac(size = 0.5) %>% 
  ggplot() + 
    geom_point(
        mapping = aes(
          x = bill_length_mm, 
          y = bill_depth_mm,
          color = species
       )
    ) + 
    xlim(30, 60) + 
    ylim(15, 22)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

checkboxGroupInput(
  inputId = "species_selection_mixed", # Think of it as an object name (not visible to users)
  label = "Penguin species:", # Title (visible to users)
  choices = c("Adelie", "Chinstrap", "Gentoo"), # Options for users to select
  selected = c("Adelie", "Chinstrap", "Gentoo") # What's selected by default
)

sliderInput(
  inputId = "sample_size_mixed", # Think of it as an object name (not visible to users)
  label = "Sample Size:", # Title (visible to users)
  min = 0, # Minimum value on slider
  max = 1, # Maximum value on slider
  value = 1, # Default value on slider
  step = 0.01 # Interval between each tick on the slider
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

renderPlot(
  penguins_df %>%
    filter(species == input$species_selection) %>% 
    sample_frac(size = input$sample_size) %>% 
    ggplot() + 
      geom_point(
          mapping = aes(
            x = bill_length_mm, 
            y = bill_depth_mm,
            color = species
         )
      ) + 
      xlim(30, 60) + 
      ylim(15, 22)
)
```
